---
title: "Predicting density dependent somatic growth in Norwegian spring spawning herring"
output: github_document
header-includes:
   - \usepackage{amsmath}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

*Contributors: Erling Kåre Stenevik<sup>1\*</sup>, Sondre Hølleland<sup>1&dagger;</sup>, Katja Enberg<sup>2</sup>, Åge Høines<sup>1</sup>, Are Salthaug<sup>1</sup>, Aril Slotte<sup>1</sup>, Sindre Vatnehol<sup>1</sup>, Sondre Aanes<sup>3</sup>*

  <sup>1</sup> Institute of Marine Research, Norway.<br>
  <sup>2</sup> University of Bergen, Norway.<br>
  <sup>3</sup> Norwegian Computing Center, Norway.<br>
  <sup>\*</sup> Corresponding author; [erling.stenevik@hi.no](emailto:erling.stenevik@hi.no)<br>
  <sup>&dagger;</sup> Responsible for the code. Correspondance related to this to: [sondre.hoelleland@hi.no](emailto:sondre.hoelleland@hi.no)
  

[Paper link goes here](https://academic.oup.com/icesjms)

This github repository contains all the code used in Stenevik et. al.(2022a). Due to regulations, not all the data is publicly available and some of the results are therefore not possible to reproduce, but we have tried to enable the user to reproduce the main results. The individual herring data (Stenevik et. al. 2022b) is published and can be downloaded either manually from the website (see reference list) or by code provided below. The XSAM time series is available in the supplementary material to the article, while the temperature data is not public. 


## Paper abstract
Density dependent growth, which might influence the effects of fisheries on a population are often ignored when management strategies are evaluated, mainly due to a lack of appropriate models readily available to be implemented. To improve on this, we investigated if somatic growth in Norwegian spring spawning herring (Clupea harengus) depend on cohort density using a formulation of the von Bertalanffy growth function on cohorts from 1921 to 2014 and found a significant negative correlation between estimated asymptotic length and density. This clearly indicates density dependent effects on growth, and we propose a model which can be used to predict size-at-age of Norwegian spring spawning herring as function of herring density (the abundance of two successive cohorts) in future estimation of reference points (FMSY) and short-term predictions of catch advice.


## Data
The main data has been put together by Stenevik et al (2022b) and is available at [https://doi.org/10.21335/NMDC-496562593](https://doi.org/10.21335/NMDC-496562593). In our *R/1_data.R* script, the individual herring data is downloaded by running the following code:
```{r download.data, echo = TRUE, eval = FALSE}
if(!("HerringData.csv" %in% list.files(path = "inputdata/") )) {
  download.file(url = "https://ftp.nmdc.no/nmdc/IMR/Herring/HerringData.csv", 
                destfile = "inputdata/HerringData.csv")
}
```

We are not at liberty to publish the XSAM series here on github, but the user can download it from the paper supplementary material (Table S3). If you save it as *inputdata/N.txt*, the *R/1_data.R* script will run as intended without adjustments to the code. We do not have permission to publish the temperature data, and these are therefore not publicly available. 

## Authors' github accounts

**Sondre Hølleland** - [holleland](https://github.com/holleland)

**Sindre Vatnehol** - [sindrevatnehol](https://github.com/sindrevatnehol)

## R version
The code has been run on the following R version.
```{r, echo = FALSE, eval = TRUE}
R.version
```

## License

The code for this project is licensed under the [GPL-3.0 License](LICENCE.md). 




[<img src="https://www.hi.no/en/hi/resources/layout/HI-logo-farger-engelsk.svg/original"
alt="Institute of Marine Research" width="100"/>](https://www.hi.no/en)

## References

Erling Kåre Stenevik (HI), Sondre Hølleland (HI), Katja Enberg (UiB), Åge Høines (HI), Are Salthaug (HI), Aril Slotte (HI), Sindre Vathehol (HI), Sondre Aanes (NR) (2022a) Predicting density dependent somatic growth in Norwegian spring spawning herring. Under review in *ICES Journal of Marine Science*. 

Erling Kåre Stenevik (HI), Sondre Hølleland (HI), Katja Enberg (UiB), Åge Høines (HI), Are Salthaug (HI), Aril Slotte (HI), Sindre Vathehol (HI), Sondre Aanes (NR) (2022b) Individual samples of Norwegian Spring Spawning herring 1935-2019 https://doi.org/10.21335/NMDC-496562593


