---
title: "Predicting density dependent somatic growth in Norwegian spring spawning herring"
output: github_document
header-includes:
   - \usepackage{amsmath}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

*Contributors: Erling Kåre Stenevik<sup>1\*</sup>, Sondre Hølleland<sup>1&dagger;</sup>, Katja Enberg<sup>2</sup>, Åge Høines<sup>1</sup>, Are Salthaug<sup>1</sup>, Aril Slotte<sup>1</sup>, Sindre Vatnehol<sup>1</sup>, Sondre Aanes<sup>3</sup>*

  <sup>1</sup> Institute of Marine Research, Norway.<br>
  <sup>2</sup> University of Bergen, Norway.<br>
  <sup>3</sup> Norwegian Computing Center, Norway.<br>
  <sup>\*</sup> Corresponding author; [erling.stenevik@hi.no](emailto:erling.stenevik@hi.no)<br>
  <sup>&dagger;</sup> Responsible for the code. Correspondance related to this to: [sondre.hoelleland@hi.no](emailto:sondre.hoelleland@hi.no)
  

[Paper reference goes here](http://www.imr.no)

In this project, we investigate how population density influences individual growth of Norwegian spring spawning herring. A parametrization of von Bertalanffy growth function (VBGF) is used to model the growth in terms of length of the fish. 

Since not all the data is available, rerunning everything is problematic. **(Note to self:  Perhaps I can make a branch which do not require the temperature data.)**

## R version
The code has been run on the following R version.
```{r, echo = TRUE, eval = TRUE}
R.version
```

## Authors' github accounts

**Sondre Hølleland** - [holleland](https://github.com/holleland)
**Sindre Vatnehol** - [sindrevatnehol](https://github.com/sindrevatnehol)

## Data
The main data is available at [link to nmd](https://nmb.hi.no). We do not have permission to publish the temperature data, and these are therefore not publicly available. 

## License

This code for this project is licensed under the [MIT licence](LICENCE.md). 

[<img src="https://www.hi.no/en/hi/resources/layout/HI-logo-farger-engelsk.svg/original"
alt="Institute of Marine Research" width="200"/>](https://www.hi.no/en)

